---
title: "Codon usage bias R analysis"
author: "Gabriel Alves Vieira"
date: "1/12/2021"
output: html_document
---

# Data exploration/transformation

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#, warning = FALSE, message = FALSE)
```

First, let's import all the libraries we'll need:

```{r libraries, results='hide'}
library('readxl') # Read Excel files
library('dplyr') # For data manipulation
library('ggplot2') # For data visualization
```

Loading codon_usage table into R:

```{r}
table_raw = read_excel("../primates/Primates_codon_usage.xlsx")[,c(1, 3:4, 8)]
#Taking only the columns we're interested in (1,3,4,8)
```

Viewing the first rows:

```{r}
head(table_raw)
```

Grouping by Aminoacid and Codon and obtaining the `mean, median, sd, min, max and 90th quantile` for each group:

```{r}
grouped_summary <- table_raw %>% 
  group_by(Aminoacid, Codon) %>%
  summarise(mean = mean(Fraction),
            median = median(Fraction),
            sd = sd(Fraction),
            min = min(Fraction),
            max = max(Fraction),
            quant90 = quantile(Fraction, probs = 0.9))
# grouping doesn't change how the data looks (apart from listing
# how it's grouped):
# However, it changes how it acts with the other dplyr verbs
head(grouped_summary) # outputting the resulting table
```

# Data visualization

## Scatterplot (mean fraction)

Scatterplot with frac_mean as dependent variable and aminoacid as independent variable:

```{r}
ggplot(grouped_summary, aes(x = as.factor(Aminoacid), y = mean, fill = Aminoacid)) + 
  geom_dotplot(binaxis = "y", stackdir = "center") +
  labs(title = "Codon usage in primates",
    subtitle = "Fraction mean per aminoacid in mitochondrial PCG's",
    caption = "Those were observed by using the mean of 199 mitochondrial coding sequences from different primate species",
    x = "Aminoacid",
    y = "Fraction")
```

Judging by this plot alone, it looks like there are more evident biases for some aminoacids, specially those with only two synonymous codons: (Gln, Glu, Lys, Met, Trp). The stop codon also has a clear bias.  Other aminoacids might have biases, but such are not that evident, since they have more synonymous codons.

Let's visualize which are the most representative codons in the more evident biases:

```{r}
grouped_summary[]
```


## Barplot (Mean fraction + standard deviation)

Barplot of mean fraction with standard deviation